---
title: "Association Rule Mining"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2024-12-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This Project is developed as an application of the Market Basket Analysis in R for recognizing word association for fake and true job postings.

Let us start by loading the packages required for the work.

```{r}
library(dplyr)
```
```{r}
library(tm)
```

```{r}
library(stringr)
```

```{r}
library(wordcloud)
library(RColorBrewer)
```

```{r}
library(arules) 
```

```{r}
library(tidyverse)

```



## Data Exploration

```{r}
true_data <- read.csv("True.csv")
fake_data <- read.csv("Fake.csv")
```

```{r}
head(true_data)
```

```{r}
columns_to_keep <- c("title", "text", "subject", "type")


fake_data<- fake_data %>% select(all_of(columns_to_keep))
colnames(fake_data) 
```


```{r}

true_data <- true_data %>% select(-date)
colnames(fake_data) 
```
Display the size of the dataset
```{r}
dim(true_data)
dim(fake_data)
```


A function clean_text for an automatic preprocessing
```{r}

library(tm)
library(SnowballC)
library(stringr)

library(SnowballC)  # For stemming
library(tm) # For stopwords
library(dplyr)

clean_text <- function(text) {
  
  # Replace email addresses with 'emailaddr' (modified pattern)
  text <- gsub('[a-zA-Z0-9\\-\\.]+@[a-zA-Z0-9]+\\.[a-zA-Z]{2,4}', 'emailaddr', text)
  
  # Replace URLs with 'httpaddr'
  text <- gsub('(http[s]?\\S+)|(\\w+\\.\\w{2,4}\\S*)', 'httpaddr', text)
  
  # Replace currency symbols with 'moneysymb'
  text <- gsub('£|\\$', 'moneysymb', text)
  
  # Replace phone numbers with 'phonenumbr'
  text <- gsub('\\+?\\(?\\d{1,2}\\)?[-\\s.]?\\d{3}[-\\s.]?\\d{3}[-\\s.]?\\d{4}', 'phonenumbr', text)
  
  # Replace numbers with 'numbr'
  text <- gsub('\\d+(\\.\\d+)?', 'numbr', text)
  
  # Remove any non-word characters (except space)
  text <- gsub('[^a-zA-Z0-9\\s]', ' ', text)
  
  # Remove extra spaces and trim leading/trailing whitespaces
  text <- gsub('\\s+', ' ', text)
  text <- gsub('^\\s+|\\s+?$', '', text)
  
  # Convert to lowercase
  text <- tolower(text)
  
  # Tokenize the text, remove stopwords, and apply stemming
  words <- strsplit(text, " ")[[1]]
  words <- words[!words %in% stopwords("en")]  # Remove stopwords
  
  # Remove single-character words (like 'a', 'b', etc.)
  words <- words[nchar(words) > 1]
  
  # Apply stemming
  words <- wordStem(words)
  
  # Return the cleaned text as a single string
  return(paste(words, collapse = " "))
}

# Apply the clean_text function to both 'title' and 'text' columns
true_data$title <- sapply(true_data$title, clean_text)
true_data$text <- sapply(true_data$text, clean_text)

fake_data$title <- sapply(fake_data$title, clean_text)
fake_data$text <- sapply(fake_data$text, clean_text)

# Combine the datasets
true_fake_job <- bind_rows(true_data, fake_data)


# Shuffle the combined dataset
shuffled_data <- true_fake_job[sample(nrow(true_fake_job)), ]

# Inspect the cleaned and shuffled data
head(shuffled_data)








```

```{r}
library(tm) 
```


```{r}
shuffled_data <- shuffled_data %>%
  mutate(
    title = title %>%
      iconv(from = "latin1", to = "UTF-8", sub = "") %>%
      str_trim() %>%
      str_to_lower() %>%
      str_remove_all("[[:punct:]]") %>%
      str_remove_all("[0-9]+") %>%
      removeWords(stopwords("en")) %>%
      str_replace_all("[^[:alnum:] ]", "") %>%
      str_replace_all("\\s+", " "),

    text = text %>%
      iconv(from = "latin1", to = "UTF-8", sub = "") %>%
      str_trim() %>%
      str_to_lower() %>%
      str_remove_all("[[:punct:]]") %>%
      str_remove_all("[0-9]+") %>%
      removeWords(stopwords("en")) %>%
      str_replace_all("[^[:alnum:] ]", "") %>%
      str_replace_all("\\s+", " "),

    subject = subject %>%
      iconv(from = "latin1", to = "UTF-8", sub = "") %>%
      str_trim() %>%
      str_to_lower() %>%
      str_remove_all("[[:punct:]]") %>%
      str_remove_all("[0-9]+") %>%
      removeWords(stopwords("en")) %>%
      str_replace_all("[^[:alnum:] ]", "") %>%
      str_replace_all("\\s+", " ")
  )

```
Combine all rows into a single text string and search for word frequency.
```{r}

library(stringr)
all_text <- str_c(shuffled_data$text, shuffled_data$subject, shuffled_data$title, collapse = " ")
# Split the text into individual words
word_list <- unlist(strsplit(all_text, " "))

# Create a frequency table of words
word_freq <- table(word_list)
word_freq <- sort(word_freq, decreasing = TRUE)


# Plot the word cloud
wordcloud(names(word_freq), freq = word_freq, min.freq = 10, 
          random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

```{r}
library(dplyr)
words_to_remove <- c("said", "president", "us","president","s","trump","will","new","states","t")

# Create pattern and remove words
pattern <- paste0("\\b(", paste(words_to_remove, collapse = "|"),")\\b")

# Apply the pattern-based removal to the specified columns
shuffled_data <- shuffled_data %>%
  mutate(
    title = str_replace_all(title, pattern, ""),
    subject = str_replace_all(subject, pattern, ""),
    text = str_replace_all(text, pattern, "")
  )

```



```{r}
# Combine columns into one
shuffled_data$combined <- paste(shuffled_data$title, shuffled_data$text, shuffled_data$subject, sep = " ")

# Process the combined column
split_combined <- strsplit(shuffled_data$combined, " ")
balanced_combined <- do.call(rbind, lapply(split_combined, function(x) { 
  length(x) <- max(lengths(split_combined))
  return(x)
}))
balanced_combined <- as.data.frame(balanced_combined)
colnames(balanced_combined) <- paste0("word", 1:ncol(balanced_combined))

# Add the label column back
balanced_combined$type <- shuffled_data$type
head(balanced_combined)
```

```{r}
# Reorder columns by moving 'type' to the first position
balanced_combined <- balanced_combined[c("type", setdiff(colnames(balanced_combined), "type"))]

# Check the new column order
colnames(balanced_combined)
```
Convert the dataframe into a transaction
```{r}
library(arules)
library(dplyr)


# Step 1: Clean and prepare the transactions list
transactions_list <- apply(balanced_combined[, -ncol(balanced_combined)], 1, function(x) x[!is.na(x) & x != ""])

# Step 2: Remove empty transactions (this ensures there are no rows with no items)
transactions_list <- transactions_list[sapply(transactions_list, length) > 0]

# Step 3: Remove duplicates within each transaction (ensure no repeated words in a single transaction)
transactions_list <- lapply(transactions_list, function(x) unique(x))

# Step 4: Optionally, sort each transaction's items (for consistency)
transactions_list <- lapply(transactions_list, function(x) sort(x))

# Step 5: Convert the cleaned list to a transaction format using 'arules'
transactions <- as(transactions_list, "transactions")

# Step 6: summary the transactions object
summary(transactions)




```

```{r}
par(las = 2)


itemFrequencyPlot(transactions, support = 0.35)  
```
```{r}
common_words = c('also','call','donald','like','peopl','presid','reuter','sai','time','year')

long_data <- balanced_combined %>%
  pivot_longer(cols = starts_with("word"), 
               names_to = "word_column", 
               values_to = "word") 
# Filter rows where "word" matches the target words
filtered_data <- long_data %>%
  filter(word %in% common_words) %>%
  group_by(word, type) %>% 
  summarise(count = n(), .groups = "drop")

# Create a bar chart
# Ensure 'type' is a factor
filtered_data$type <- as.factor(filtered_data$type)

# Create the plot
# Calculate the percentage for each word within its category
filtered_data <- filtered_data %>%
  group_by(word) %>%
  mutate(percentage = (count / sum(count)) * 100)

# Create the plot with percentages
ggplot(filtered_data, aes(x = reorder(word, -percentage), y = percentage, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Percentage of Words in True vs Fake News",
    x = "Word",
    y = "Percentage (%)",
    fill = "News Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  scale_fill_brewer(palette = "Set1")



```
Performing the apriori algorithm
```{r}
textrules <- apriori(transactions, parameter = list(support =
                                                      0.01, confidence = 0.25, minlen = 2))
```

```{r}
textrules
```
Evaluating Model Performance
```{r}
summary(textrules)
```

```{r}
library(arules)
inspect(head(textrules, 20))
```

